\section{Hacker's Delight}\label{sec:book}
\epigraph{
``There is nothing here about hacking
in the sense of breaking into computers.''
}{
\emph{``Hacker's Delight''} Website
\cite{Warren:HD:Website}
}


\subsection*{The Book}
Henry Warren's book ``\emph{Hacker's Delight}''
is described by himself as
``a collection of small programming tricks''
\cite{Warren:2012:HD:2462741} (p. \texttt{xv}).
In fact, the table of contents reads like
an encyclopedia of formulas for bit level manipulation:
From detecting, counting and changing specific single bits
to formulas that fulfil a mathematical goal
(and interpret the bitword as a number, not a vector),
everything imaginable seems to be part of this collection.


\subsection*{This Paper}
Inspired by and named after ``\emph{Hacker's Delight}'',
this paper complements it with methods and explanations
that have been left out by the author (intentionally):

\begin{quote}
``The presentation is informal.
Proofs are given only when the algorithm is not obvious,
and sometimes not even then.''
\par\hfill -- \emph{Hacker's Delight},
page \texttt{xv} \cite{Warren:2012:HD:2462741}
\end{quote}

Instead of simply explaining given formulas,
this essay assembles them from a few elementary operations
in a stepwise manner.
Besides a fundamental understanding of the covered topics,
the aim is to provide the reader with methods
that can be adapted and thereby used to create new formulas.


\subsection*{The Book's Title}
One of the disclaimers given at multiple occasions by Henry Warren
considers the possibility of mistaking ``Hacker''
for someone breaking into others computers.
Rather, the two words forming the book's title
can be described as follows:
\begin{description}
\item[Hacker]
Someone who enjoys clever ways of coding
and isn't necessarily a professional.
The created work might be useful
or can just be seen as the result of a clever game.

\item[Delight]
Once a certain liking for these kind of problems has been developed,
the feeling when understanding elegant solutions is pure joy.
\end{description}
